The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-03-12 paper/iphas-dr2.tex
Updated caption to Fig 13

In contrast, 
closer towards the galactic centre at $l=45^{\circ}$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
In the colour magnitude diagram it is clear that the stars are split into two distinct groups,
with one significantly redder than the other.
The bluer group is composed of main sequence stars,
with the slope of this group in the colour magnitude diagram attributable to the significantly increasing extinction.
Meanwhile the redder group is principally composed of red giant stars.
As these stars are intrinsicly brighter,
at some given apparent magnitude they will be substantially further away than their main sequence counterparts.
Given that, along this sightline,
extinction continues to increase with distance,
the red giants we observe will be subject to more reddening than a main sequence star exhibiting the same apparent magnitude.
Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequnce and giant branch stars.

Finally, in one of our earliest sightlines at $l=30^{\circ}$
we find a very high number of extremely reddened giants
in addition to an unreddened population
of foreground dwarfs.
In contrast to the sightline at $l=45^{\circ}$,
there is no clear group of giant stars visible in the colour magnitude diagram of Fig.~\ref{fig:l30}.
Though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibited a relativly narrow range of reddenings as the lay beyond most or essentially all of the Galactic dust.
At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
even at the substantial distances at which we can observe reddened giant stars extinction is continuing to rise and we still find ourselves well within the Galactic dust disc.
It is also apparent that the $(r-i)$ width of both the M dawrfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
This is indicative of a steeper rise in reddening,
both within several hundred parsecs (M dwarfs)
and within a few kiloparsecs (early A dwarfs).

@@

>>>>

In contrast, 
closer towards the galactic centre at $l=45^{\circ}$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
In the colour-magnitude diagram it is clear that the stars are split into two distinct groups,
with one significantly redder than the other.
The bluer group is composed of main sequence stars,
with the slope of this group in the colour-magnitude diagram attributable to the significantly increasing extinction.
Meanwhile the redder group is principally composed of red giant stars.
As these stars are intrinsicly brighter,
at some given apparent magnitude they will be substantially further away than their main sequence counterparts.
Given that, along this sightline,
extinction continues to increase with distance,
the red giants we observe will be subject to more reddening than a main sequence star exhibiting the same apparent magnitude.
Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequnce and giant branch stars.

Finally, in one of our earliest sightlines at $l=30^{\circ}$
we find a very high number of extremely reddened giants
in addition to an unreddened population
of foreground dwarfs.
In contrast to the sightline at $l=45^{\circ}$,
there is no clear group of giant stars visible in the colour-magnitude diagram of Fig.~\ref{fig:l30},
though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibit a relativly narrow range of reddenings as the lie beyond most or essentially all of the Galactic dust.
At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
even at the substantial distances at which we can observe reddened giant stars extinction is continuing to rise and we still find ourselves well within the Galactic dust disc.
It is also apparent that the $(r-i)$ width of both the M dawrfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
This is indicative of a steeper rise in reddening,
both within several hundred parsecs (M dwarfs)
and within a few kiloparsecs (early A dwarfs).

@@

\begin{figure*}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=0.5\textwidth]{./plots/ccd-180-3.pdf} 
        \includegraphics[width=0.5\textwidth]{./plots/cmd-180-3.pdf}
    \end{minipage}
    \caption{Colour-colour and colour-magnitude diagram (left and right panel)
    showing sources flagged as \emph{veryReliable}
    located in an area of one square degree
    centred near the Galactic anti-centre 
    at $(l,b)=(180^\circ,+3^\circ)$.
    The colour-colour diagram shows the
    position of the main sequence (thin solid line),
    giant stars (thick solid line)
    and the reddening track for an A0V-type star (dashed line)
    based on the \citet{Pickles1998} library of empirical spectra.
    The colour-magnitude only show the reddening vector
    along with the unreddened 1~Gyr isochrone due to \citet{Bressan2012},
    which has been placed at an arbitrary distance of 2~kpc for reference.
    This is one of the least reddened sightlines
    in the survey %(Av XX, Schlegel et al. XXXX)
    and hence the observed stellar population appears to be dominated 
    by lowly reddened main sequence stars.}
    \label{fig:l180}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=0.5\textwidth]{./plots/ccd-45-2.pdf}
        \includegraphics[width=0.5\textwidth]{./plots/cmd-45-2.pdf}
    \end{minipage}
    \caption{Same as above for $(l,b)=(45^\circ,+2^\circ)$,
    which is one of the highest-density sightlines in the survey,
    revealing two groups of stars in colour-magnitude space.}
    \label{fig:l45}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=0.5\textwidth]{./plots/ccd-30-0.pdf}
        \includegraphics[width=0.5\textwidth]{./plots/cmd-30-0.pdf} 
    \end{minipage}
    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$.
    This is one of the most reddened sightlines in the survey.
    %(XXX, Schlegel et al XXXX).
    }
    \label{fig:l30}
\end{figure*}
